### 2024-02-24 dissertation-jhu/06-chapter-2.tex
published repository

\chapter{Chapter title goes here} \label{chap:chap-2}

% epigraph before chapter heading
\epigraphhead[0]{\epigraph{\enquote{Do not believe everything you see on the internet}}{-- Albert Einstein}}

\blindtext