### 2024-03-04 dissertation-jhu/06-chapter-2.tex
updated short-header format

\chapter{Chapter title goes here} \label{chap:chap-2}

% if you want a short header you can use the following command
% \chapter[short-header-name]{chapter-title} \label{chap:chap-2}

% epigraph before chapter heading
\epigraphhead[0]{\epigraph{\enquote{Do not believe everything you see on the internet}}{-- Albert Einstein}}

% remove the following and add your chapter text
\blindtext

\begin{figure}[ht]
\begin{center}
    \includegraphics[width=\textwidth, trim={6cm 5cm 6cm 5cm},clip,page=1] {chap2.pdf}
    \caption{Here are some photos of ducks to make you feel happy in tough times.}
    \label{fig:ducks}
\end{center}
\end{figure}

\blindtext[2]       % generates a single paragraph of double size