### 2024-03-18 beamer/theme-preamble.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% JH colors %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% following colors are defined based on: https://brand.jhu.edu/color
\definecolor{HeritageBlue}{RGB}{0,45,114}       % heritage blue
\definecolor{SpiritBlue}{RGB}{114,172,229}      % spirit blue

\definecolor{JHUPastle}{RGB}{0,155,119}         % pastle
\definecolor{JHUDeepBlue}{RGB}{0,114,206}       % deep blue

\definecolor{JHURed}{RGB}{166,25,46}            % red
\definecolor{JHUGreen}{RGB}{40,97,64}           % green

\definecolor{JHUViolet}{RGB}{81,40,79}          % violet
\definecolor{JHUPurple}{RGB}{161,90,149}        % purple
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% JH COLORS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%% COLOR PALETTE SETTINGS %%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% JHU color is used to set the palette for the beamer template

\setbeamercolor{palette primary}{bg=HeritageBlue,fg=white}
\setbeamercolor{palette secondary}{bg=HeritageBlue,fg=white}
\setbeamercolor{palette tertiary}{bg=HeritageBlue,fg=white}
\setbeamercolor{palette quaternary}{bg=HeritageBlue,fg=white}
\setbeamercolor{sidebar}{bg=White}
\setbeamercolor{palette sidebar primary}{bg=white,fg=HeritageBlue}
\setbeamercolor{palette sidebar secondary}{bg=white,fg=HeritageBlue}
\setbeamercolor{palette sidebar tertiary}{bg=white,fg=HeritageBlue}
\setbeamercolor{palette sidebar quaternary}{bg=white,fg=HeritageBlue}

\setbeamercolor{titlelike}{bg=HeritageBlue,fg=White}
\setbeamercolor{author}{fg=HeritageBlue}
\setbeamercolor{institute}{fg=HeritageBlue}
\setbeamercolor{date}{fg=HeritageBlue}

\setbeamercolor{frametitle}{bg=HeritageBlue,fg=White}
\setbeamercolor{frametitle continuation}{bg=HeritageBlue,fg=White}

\setbeamercolor{section in toc}{fg=HeritageBlue}        % TOC sections
\setbeamercolor{structure}{fg=HeritageBlue}             % itemize, enumerate, etc

\setbeamercolor{title in sidebar}{fg=HeritageBlue}
\setbeamercolor{author in sidebar}{fg=SpiritBlue}

\setbeamercolor{block title}{bg=HeritageBlue,fg=White}
\setbeamercolor{block body}{bg=HeritageBlue!10,fg=Black} 
\setbeamercolor{block title alerted}{bg=JHURed,fg=White}
\setbeamercolor{block body alerted}{bg=JHURed!10,fg=Black} 
\setbeamercolor{block title example}{bg=JHUGreen,fg=White}
\setbeamercolor{block body example}{bg=JHUGreen!10,fg=Black}

\setbeamercolor{footnote}{bg=White,fg=HeritageBlue}
\setbeamercolor{footnote mark}{bg=White,fg=HeritageBlue}

\setbeamercolor{alerted text}{fg=JHURed}

%%%%%%%%%%%%%%%%%%% END COLOR PALETTE SETTINGS %%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%% FONT SETTINGS %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% size of the fonts in different environments of the beamer template
\setbeamerfont{title}{size=\TitleFontSize}
\setbeamerfont{subtitle}{size=\SubtitleFontSize}
\setbeamerfont{author}{size=\AuthorFontSize}
\setbeamerfont{institute}{size=\InstituteFontSize}
\setbeamerfont{date}{size=\DateFontSize}

\setbeamerfont{frametitle}{size=\FrameTitleFontSize}

%%%%%%%%%%%%%%%%%%%%%%%%%% END FONT SETTINGS %%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%% FRAME TITLE, SIDEBAR, FOOTER SETTINGS %%%%%%%%%%%%%%%%%
\makeatletter
    \logo{\includegraphics[height=\FrameHeadHeight]{JHU_shield.pdf}}
    \setlength{\beamer@headheight}{\FrameHeadHeight}
\makeatother

\makeatletter
%%% following two lines removes the slide title and author from the sidebar
    \setlength\beamer@sidebarwidth{\SidebarWidth}
    \setbeamertemplate{sidebar \beamer@sidebarside} 
  { %%% following code snippet customizes the width of the sidebar
    \beamer@tempdim=\beamer@sidebarwidth%
    \advance\beamer@tempdim by -6pt%
    \insertverticalnavigation{\beamer@sidebarwidth}%
    \vfill
    \ifx\beamer@sidebarside\beamer@lefttext%
    \else%
      \usebeamercolor{normal text}%
      \llap{\usebeamertemplate***{navigation symbols}\hskip0.1cm}%
      \vskip2pt%
    \fi%
  }%
\makeatother

%%% following code snippets removes the navigation symbols 
%%% and inserts the frame number/total frame in the footer
\setbeamertemplate{navigation symbols}
{\tiny\insertframenumber/\inserttotalframenumber}
%%%% TO-DO: MORE CUSTOMIZATION OF THE FOOTER

%%%%%%%%%%%%% END FRAME TITLE, SIDEBAR, FOOTER SETTINGS %%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%% TOC SETTINGS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% the next block of commands puts the table of contents at the 
%%%% beginning of each section and highlights the current section
\AtBeginSection[]
{ %%% use \tableofcontents[hideothersection] for different outline appearance
  \begin{frame}
    \frametitle{Outline}
    \tableofcontents[currentsection]
  \end{frame}
}

%%%%%%%%%%%%%%%%%%%%%%%%% END TOC SETTINGS %%%%%%%%%%%%%%%%%%%%%%%%%%%%